---
title: "W11 Exercise - Poisson Regression"
author: "Xingyu Hu"
format: 
    html:
        theme: default
        embed-resources: true
        code-block-bg: true
        comments: 
            hypothesis: true
        toc: true
        toc-location: left
        number-sections: true
---

As scientist studies the effect of seizure drugs on the number of seizures during the first eight weeks of treatment. The dataset is available in the **robust** package:

```{r}
data("breslow.dat", package = "robust")
```

Carry out a Poisson regression of the number of seizures after treatment (`sumY`) as the dependent variable, and the treatment conditions (`Trt`), `Age`, and the number of seizures before treatment (`Base`) as the independent variable.

```{r}
fit.breslow <- glm(sumY ~ Trt + Age + Base, 
                   data = breslow.dat, 
                   family = poisson())
summary(fit.breslow)
exp(coef(fit.breslow))
stargazer::stargazer(fit.breslow, type = 'text')

fit.breslow.od <-
  glm(sumY ~ Trt + Age + Base,
      data = breslow.dat,
      family = quasipoisson())

summary(fit.breslow.od)
```

```{r}
#| warning: false

fit.breslow.dat <- glm(sumY ~ Base+Age+Trt, data=breslow.dat, family=poisson())
summary(fit.breslow.dat)

fit.breslow.dat.reduced <- glm(sumY ~ Base+Age+Trt, data=breslow.dat, family=poisson())
summary(fit.breslow.dat.reduced)

exp(coef(fit.breslow.dat.reduced))

equatiomatic::extract_eq(fit.breslow.dat.reduced, use_coefs = TRUE)
stargazer::stargazer(fit.breslow.dat.reduced, type = 'text')

library(qcc)
qcc.overdispersion.test(breslow.dat$sumY, type = "poisson")

fit.breslow.dat.reduced.od <- glm(sumY ~ Base+Age+Trt, data=breslow.dat, family = quasipoisson())
summary(fit.breslow.dat.reduced.od)
```


