---
title: "W4 Exercise - Characters"
author: "Xingyu.Hu"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: hpstr
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(dplyr)
library(tidyr)
library(DescTools)

setwd("D:/Documents/RProj/ENV222_codes")
student_names <- read.csv("data/student_names.csv")
```



1. Here is a sentence in German:

```{r}
x <- 'Victor jagt zwölf Boxkämpfer quer über den großen Sylter Deich'
```


a. How many characters are there in the sentence, excluding the blank spaces?

```{r}
num <- gsub(" ", "", x) %>%
   nchar() %>% 
    print()
```

b. Do these characters cover all the German letters?

```{r}
german_letters <- letters %>% 
    append(c("ä", "ö", "ü", "ß")) %>% 
    print()

strsingle <- tolower(x) %>%
    gsub(" ", "", .) %>%
    strsplit("") %>%
    print()

check_german <- function(input_str){
    for (i in 1:length(german_letters)) {
        if (grepl(german_letters[i], input_str) == FALSE){
            return("Input string does not contain all German letters!")
        }
    }
    return("Input string contains all German letters!")
}

check_german(strsingle)
```

c. Which character is repeated the most times?

```{r}
word_freq <- table(strsingle) %>% 
    data.frame() %>%
    arrange(desc(Freq))

head(word_freq, 10)
```

d. Which is the longest word in the sentence?

```{r}
strword <- strsplit(x, " ") %>% 
    data.frame()
names(strword) <- "word"

strword$word[which.max(nchar(strword$word))]
```

e. Give a new sentence with the number replaced with Arabic number (zwölf -> 12).

```{r}
print(gsub("zwölf", "12", x))
```

---

2. Process the *student_name* dataset.

a. Whose name is the shortest?

```{r}
student_names$Name[which.min(nchar(student_names$Name))]
```

b. Add a new column as the family name. How many family names do we have? Which family name is repeated the most times?

```{r}
student_names$FamilyName <- student_names$Name %>%
    strsplit(" ") %>%
    sapply(tail, 1)

head(student_names, 5)
tail(student_names, 5)

name_freq <- table(student_names$FamilyName) %>% 
    data.frame() %>%
    arrange(desc(Freq))

head(name_freq, 10)
```

c. Add a new column with the family name followed by the given name.

```{r}
names <- student_names %>%
    separate(col = Name,
            into = c("GivenName", "FamilyName"),
            sep = " ")

student_names$NewName <- paste(names$FamilyName, names$GivenName)

head(student_names, 5)
tail(student_names, 5)
```

d. Order the data frame by the family name.

```{r}
student_names <- student_names[order(student_names$FamilyName),]

head(student_names, 10)
tail(student_names, 10)
```

---

3. Text could be encrypted in a simple rule like this: C $\rightarrow$ D, A $\rightarrow$ B, K $\rightarrow$ L, E $\rightarrow$ F: CAKE $\rightarrow$ DBLF(+1); D $\rightarrow$ G, A $\rightarrow$ E, K $\rightarrow$ O, E $\rightarrow$ I: CAKE $\rightarrow$ GEOI(+4). Decrypt the following sentence which was encrypted with the rule above:

    > skt grcgey xkskshkx rubk hkigayk ul xusgtik utre
    
```{r}
encrypteIt <- function(input, move = 1){
    if(move > 25 | move < -25){
        return("Please set the move between -25 and 25")
    }


    asciiSingle <- strsplit(input, "") %>%
        data.frame() %>%
        apply(2, DescTools::CharToAsc)

    for(i in 1:length(asciiSingle)){
        if(asciiSingle[i] == 32){
            next
        }

        if(asciiSingle[i] + move > 122){
            asciiSingle[i] <- asciiSingle[i] + move - 26
            next
        }

        if(asciiSingle[i] + move < 97){
            asciiSingle[i] <- asciiSingle[i] + move + 26
            next
        }

    asciiSingle[i] <- asciiSingle[i] + move
    }

    output <- paste(apply(asciiSingle, 2, DescTools::AscToChar))
    return(output)
}

text <- "skt grcgey xkskshkx rubk hkigayk ul xusgtik utre"
for(i in 1:25){
    print(paste0(encrypteIt(text, i), " (+", i, ")"))
}
```