---
title: "W8 Exercise - MANOVA"
author: "Xingyu Hu"
date: "`r Sys.Date()`"
format: 
    html:
        theme: default
        embed-resources: true
        code-block-bg: true
        comments: 
            hypothesis: true
        toc: true
        toc-location: left
        number-sections: true
---

# `iris` dataset
Do the species have influence on the petal size? Carry out MANOVA with the post-hoc test and give your conclusion.

```{r}
#| warning: false
#| message: false

library(ggplot2)
library(tidyr)
iris[, c("Petal.Length", "Petal.Width", "Species")] |> 
    pivot_longer(cols = c(Petal.Length, Petal.Width)) |> 
    ggplot() +
    geom_boxplot(aes(Species, value, fill = name), outlier.size = 0.8) +
    labs(y = "Size (cm)", fill = " ")

library(gplots)
par(mfrow = c(1, 2))
plotmeans(iris$Petal.Length ~ iris$Species)
plotmeans(iris$Petal.Width ~ iris$Species)
```
## Hypothesis

- $H_0$: The population means of petal length and the petal width are not different across the species.
    
## Collect data

```{r}
# multivariate normality test
PetalSize <- cbind(iris$Petal.Length, iris$Sepal.Width)
```

## Calculate the test statistic: the Pillai's Trace test or the F-test

```{r}
iris_manova <- manova(PetalSize ~ iris$Species)
summary(iris_manova)
summary(iris_manova, test = "Pillai")
summary(iris_manova, test = "Wilks")
summary(iris_manova, test = "Roy")
summary(iris_manova, test = "Hotelling-Lawley")
```

Univariate ANOVAs for each dependent variable:

```{r}
summary.aov(iris_manova)
```

**Conclusion:** The species has a statistically significant effect on the petal width and petal length.

## Post-hoc test

**Linear Discriminant Analysis (LDA)**

```{r}
library(MASS)
iris_lda <- lda(iris$Species ~ PetalSize, CV = FALSE) |> print()

plot_lda <- data.frame(Species = iris$Species,
                       lda = predict(iris_lda)$x)

ggplot(plot_lda) +
    geom_point(aes(x = lda.LD1, y = lda.LD2, colour = Species))
```

**Conclusion:** The petal size of the setosa species is different from other species.